---
title: "Cooperative Learning"
author: "Leandro Kovalevski"
toc: true
number-sections: true
toc-depth: 2
format:
  gfm:
    code-fold: true
    code-summary: "Show the code"
theme:
  light: zephyr
  dark: cyborg
---


#  Executive summary {-}

- to be completed..
- ... 
- ...





# Settings {-}

```{r}
file_name <- "df_bcra.rds"
```
 
This report was run with the **objetive** of describing the bcra dataset: \n


 -  **Run date**: `r  Sys.Date()`
 
 -  **Dataset**: `r  file_name` 




# Dataset description.

The database consists of a random sample of 23,857 tax identification numbers (CUITs) belonging to individuals who had at least one debt in the Argentine financial system in June 2019, and were in credit situation 1 or 2 (meaning they did not have overdue payments exceeding 90 days), obtained from the debtor database provided by the Central Bank of the Argentine Republic (BCRA) for the period of June 2019.
For the tax identification numbers in the random sample, debts in all entities were recorded and summarized for June 2019, as well as for the previous 6 months. Debts of these tax identification numbers between July 2019 and June 2020 were also recorded to assess their evolution. 
The response variable is a binary variable constructed from the most severe credit situation of the tax identification number (CUIT) between the periods of July 2019 and June 2020. The variable takes the value 1 if the most severe credit situation is greater than or equal to 3 in any debt any period, and 0 otherwise.
In the dataset 'df_bcra.rds', the information recorded with 28 variables is available. The data is anonymized and variable names are not displayed.

```{r}
#' ## Load data and needed packages.

#' Install (if needed)  'here' package to use relative paths. 
if(!("here" %in% installed.packages()[, "Package"])){ 
  install.packages("here") 
  }

#' Load generic functions ----
source(here::here("src", "utils.R"), encoding = "UTF-8")

#' Cargar las librerías necesarias
loadPackages(c(
  "here", "multiview", "ggplot2", "knitr", "scales", "dplyr", "doBy", "moments",
  "gains", "ROCR", "skimr", "moments", "corrplot"
  ))

#' Set data path
file_path <- here::here("data", "ready")

```


